---
title: "Appendix Dogs and MOT sensitivity to outliers and transformation  "
author: "Ian Philips"
date: "06/08/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = F)
```


# Introduction

The first part of the appendix reruns the analysis with outliers removed 


The second part of the appendix reruns the analysis with outliers removed AND log transformation of dogs per person.  

The finding is that in both analyses dogs per person  is a variable associated with car use.  Dogs per person continues to be a predictor of car km travelled even even when other covariates are considered.  





## Analysis with some outliers removed  

this is an appendix showing a run of the analysis where I have removed outliers.  I set outliers as the top 1% of postcode outcodes in terms of dogs per person.  

This means I removed 22 post code outcodes.  Each of those has >= 0.9580356 dogs per person.  


When I ran the analysis with these outliers removed the same finding was reached.  There is still a strong association between car km travelled per person and dog ownership per person.  With outliers removed, this association also still holds when other covatiates are considered.  




```{r}
#packages

#these are the main ones for wrangling
# a few others are found in later chunks 
library(tidyverse)
library(tmap)
library(sf)

```


First read in the data sets to use.  There is a table detailing these in the paper.  

```{r}
#--------------read datasets to be wrabngled ------------------------
#dog data 
dogs <- read_csv("data/DFEFAdogdata.csv")
postcode_oa <- read_csv("data/OA_postcode.csv")
lookup <- read_csv("data/OA11_LSOA11_MSOA11_LAD11_EW_LUv2.csv")

postcodepop <- read_csv("population_postcodes/Postcode_Estimates_Table_1.csv")
sum(postcodepop$Total)

#This is a set of various variables including rural urban classification
clusterdata <- read_csv("MOTdata/LSOAamasterEcc030418subset.csv")
clusterdata <- clusterdata %>% select(LSOA11CD,income,green900,RUC11,ru5)


#MOT data (version 9)
mot <- read_csv("data/v9carsAndPop.csv")


#spatial data 
#lsoa polygons
lsoa <- st_read(dsn = "LSOA11SG", layer = "LSOAEW11SG")


#english and welsh LSOA population weighted centroids
pwc <- st_read("data/dist_to_stations.gpkg") # pop weighted cents with rail access info/  
pwc <- pwc %>% select(code,name,dist_to_nearest_station)



#geodemographic data 
income <- read_csv("data/ExperianMedianHHIncome2011.csv")
census <- read_csv("data/selected_UKcensus2011.csv")
census <- census %>% select(LSOA11CD,nocar_percent,
                            car1_percent,
                            car2_percent,
                            car3_percent,
                            car4_or_more_percent,
                            age_median,
                            white_percent,
                            dwel_detached_percent,
                            dwel_semi_detached_percent,
                            dwel_terraced_percent,
                            flat_percent,
                            econ_all_econ_active_percent)


female <- read_csv("data/urp_pc_female.csv")
female <- female %>% select(llsoa11cd,pc_female)

census <- left_join(census,female,by = c("LSOA11CD" = "llsoa11cd" ))


#vul index I didn't use this in tha analysis in the end 
vul <- read_csv("data/Mattiolietal.csv")


#postcode districts 
pcdist <- st_read("GB_Postcodes/PostalDistrict.shp")
#qtm(pcdist)

#summary(census$pc_female)

```

##  Wrangle the data

The spatial resolution for analysis is Postcode outcode.    

Car use data is available at LSOAa resolution and dog ownership is at a coarser non-overlapping postcode outcode geography (e.g. LS21).  To avoid ecological fallacy (Openshaw, 1984)  we aggregated to postcode outcode resolution .  



```{r}
#wrangle data

##### wrangling data ###################

#We have data at 2 spatial resolutions, dog data at postcode district resolution which are larger areas
#than lsoas which is the resolution of the MOT data.  
#WE can aggregate to postcode districts: this smooths out some vatiation in the mot vehicle use data
#or we can attribute the dog ownership level to each LSOA.  This preseerves variation in MOT vehicle use data,
#but there can be some ecological fallacy issues.  




#----------- wrangle the dogs data so the spatial resolution is postcode districts.   -----------

#wrangle to postcodes to outcode (e.g. LS2 the postcode district)
#generate a human population estimate for the postcode districts, by joining dta to the dogs data 
#spatial join the postcode polygon attributes to points 
#join the dog data.  


#------------------ join postcode population to postcode district ----

#first aggregate postcode population to postcode district population 
postcode_split2 <- postcodepop %>%
  extract(Postcode, into = c('out', 'in'), '(\\S{2,4})\\s*(\\d\\w\\w)') %>%
  mutate(Postcode = sprintf('% -4s%s', out, `in`))

#names(postcodepop)
#names(postcode_split2)

#group 
groups <- group_by(postcode_split2,out)
#summarize
postcodedistpop <- dplyr::summarize(groups, totalhumanpop = sum(Total))
#sum(postcodedistpop$totalhumanpop)


#this gives a population per postcode district.  Next we can 
#make a postcode area variable of dogs per person.  
#this will eventually get compared to the MOT miles per person variable. 

dogs <- left_join(dogs,postcodedistpop,by = c("PostcodeDistrict" = "out"))
names(dogs)
#sum(dogs$totalhumanpop,na.rm=T)
#sum(dogs$EstimatedDogPopulation,na.rm=T)

dogs$Dogs_pp <- dogs$EstimatedDogPopulation / dogs$totalhumanpop
#summary(dogs$Dogs_pp)
#summary(dogs$totalhumanpop)

#plot(dogs$totalhumanpop,dogs$EstimatedDogPopulation,main = "Dog population vs Human Population")

#cor(dogs$totalhumanpop,dogs$EstimatedDogPopulation,use= "pairwise.complete.obs",method = "pearson")
#0.437772
#cor(dogs$totalhumanpop,dogs$EstimatedDogPopulation,use= "pairwise.complete.obs",method = "spearman")
#0.5213059
#there is a moderately strong correlation, th plot shows a cone.  
#There is likely another factor / factors which might provide a stronger 
#relationship with multiple regression 


#names(dogs)
#[1] "PostcodeDistrict"       "EstimatedDogPopulation" "totalhumanpop"          "Dogs_pp"  

n = dogs %>% select(PostcodeDistrict) %>% distinct()
duplicates = dogs %>% 
  group_by(PostcodeDistrict) %>% 
  filter(n()>1)



#------------- process LSOA data ------------------ 

#you can uncomment this and make a map if you want but it takes a while to render
# tm_shape(pcdist)+
#   tm_polygons(alpha = 0.5)+
#   tm_shape(pwc)+
#   tm_dots(col = "red")

#------------ select columns required from the mot data
#names(mot)
mot2 <- mot %>% select(LSOA,Pop,HH,N,km_pp,km_perAdult,emfac_co2Av,litresAv,sizeAv)


pwc <- left_join(pwc,mot2,by = c("code" = "LSOA"))
pwc <- pwc %>% filter(!is.na(km_pp)) # removes scottish lsoa/dz as we only have dogs data for England and Wales 


pwc<- left_join(pwc,clusterdata,by = c("code" = "LSOA11CD"))
pwc<- left_join(pwc,income,by = c("code" = "LSOA11CD"))
#make an income quintile
pwc <- pwc %>% mutate(income_quintile = ntile(Median_Household_Income, 5))

#remove outlier areas where mean kmpp is over 20,000 pa 
filtered <- pwc %>% filter(km_pp > 20000)
pwc <- pwc %>% filter(km_pp < 20000)


#join selected census variables 
pwc<- left_join(pwc,census,by = c("code" = "LSOA11CD"))
pwc$multicarpc <- pwc$car4_or_more_percent + pwc$car3_percent +pwc$car2_percent


#join vul index
pwc <- left_join(pwc,vul, by = c("code" = "LSOA11CD"))

#######---------- spatial joining -------########


#----- make data at postcode district resolution ---------
#st_crs(pwc)
pcdist <- st_transform(pcdist,crs = 27700)

#names(pwc)

pcdist2 <- st_join(pcdist, pwc %>% filter(!is.na(ru5))) %>% group_by(PostDist) %>% 
  summarise(meankm_pp = mean(km_pp,na.rm = T),
            meankm_per_adult = mean(km_perAdult,na.rm = T),
            mean_emfac_co2Av = mean(emfac_co2Av,na.rm = T),
            mean_engine_size = mean(sizeAv,na.rm = T),
            mean_fuelpp_l = mean(litresAv,na.rm = T),
            tot_pop_mot = sum(Pop,na.rm = T),
            num_cars = sum(N,na.rm = T),
            HH_mot = sum(HH,na.rm = T),
            dist_to_nearest_station = mean(dist_to_nearest_station,na.rm = T),
            median_hh_income = mean(Median_Household_Income,na.rm = T),
            ru5 = min(ru5,na.rm = T),
            nocar_percent = mean(nocar_percent,na.rm = T),
            multicarpc = mean(multicarpc,na.rm = T),
            age_median = mean(age_median,na.rm = T),
            white_percent = mean(white_percent,na.rm = T),
            dwel_detached_percent = mean(dwel_detached_percent,na.rm = T),
            dwel_semi_detached_percent = mean(dwel_semi_detached_percent,na.rm = T),
            dwel_terraced_percent = mean(dwel_terraced_percent,na.rm = T),
            flat_percent = mean(flat_percent,na.rm =T),
            econ_all_econ_active_percent = mean(econ_all_econ_active_percent,na.rm =T),
            vul = mean(vul,na.rm = T),
            pc_female = mean(pc_female,na.rm = T)
            )

#---remove scotland unfortunately because *data* --
#scotland postcodes will have na for km _pp because the MOT data I read in is for England and Wales
pcdist2 <- pcdist2 %>% filter(!is.na(meankm_pp))

#---  deal with RU5 
#summary(pwc$ru5)
#summary(pcdist2$ru5)


test <- pcdist2 %>% filter(is.na(ru5)) 

pcdist2$ru5 <- as.factor(pcdist2$ru5)

pcdist2 <- pcdist2 %>% mutate(income_quintile = ntile(median_hh_income, 5))
#summary(pcdist2$income_quintile)

#--------- check for duplicates ---- 
#duplicates = pcdist2 %>% 
#  group_by(PostDist) %>% 
#  filter(n()>1) #no duplicates found


#note outliers of a mean of over 20000km per person pa per lsoa have been removed
#remove 9 outlier postcode districts #done above
#filtered <- pcdist2 %>% filter(meankm_pp > 20000) #done above

#hist(pcdist2$meankm_pp)

#map the mot data mean km per person aggregated to postcode districts
#tm_shape(pcdist2)+
#  tm_polygons(col = "meankm_pp",n=5,style = "quantile",
#              alpha = 0.5)

#glimpse(pcdist2)


#-----join to the dogs data ------
pcdist2 <- left_join(pcdist2,dogs,by = c("PostDist" = "PostcodeDistrict" ))

#there are no missing values for dog population 
pcdist2 %>% filter(is.na(EstimatedDogPopulation))

#make the dogs per person metrics with both population figures
pcdist2$Dogs_ppmot <- pcdist2$EstimatedDogPopulation/pcdist2$tot_pop_mot
pcdist2$Dogs_pppcd <- pcdist2$EstimatedDogPopulation/pcdist2$totalhumanpop

#base r plot
#plot(pcdist2$Dogs_ppmot,pcdist2$Dogs_pppcd)
#ggplot
# pcdist2 %>% 
#   ggplot()+
#   geom_point(aes(x = Dogs_ppmot, y = Dogs_pppcd))+
#   geom_smooth(aes(x = Dogs_ppmot, y = Dogs_pppcd))


#cor(pcdist2$Dogs_ppmot,pcdist2$Dogs_pppcd,use= "pairwise.complete.obs",method = "pearson")
#0.96
#cor(pcdist2$Dogs_ppmot,pcdist2$Dogs_pppcd,use= "pairwise.complete.obs",method = "spearman")
#0.99

#dissimilarity increases with high values.  the outliers are for places 
#with more than 1 dog per person mot and more than 1.5 pcd
#summary(pcdist2$Dogs_ppmot)
#summary(pcdist2$Dogs_pppcd)

#---- human population estimates seem consistent
#we have an estimate of postcode population, and also census population from LSOAs.  
#when lsoa data is aggregated to postcode level it seems consistent 
#correlation between human population estimates is very high and linear
# pcdist2 %>% 
#   ggplot()+
#   geom_point(aes(x = tot_pop_mot, y = totalhumanpop))+
#   geom_smooth(aes(x = tot_pop_mot, y = totalhumanpop))


#4 postcode districts have more than 1.5 dogs per person, 
#DL11, Upper Swaledale 
#LA20 duddoneskdale and wasdale,
#LL23, Bala N Wales 
#and 
#PO41 Isle of white Yarmouth 
topdogs <- filter(pcdist2, Dogs_ppmot >1.5)



#use dogs postcode district in the analysis 
pcdist2$Dogs_pp <- pcdist2$Dogs_pppcd

#------ map the dogs data --------------
# #tm_shape(pcdist2)+
# #  tm_polygons(col = "Dogs_pp",n=5,style = "quantile",
# #              alpha = 0.5)
# 
# tm_shape(pcdist2)+
#   tm_polygons(col = "Dogs_pp",n=5,style = "fisher",
#               alpha = 0.5)


pcdist2$cars_pp_pcdist <- pcdist2$num_cars/pcdist2$totalhumanpop





```


```{r}

#---------save key datasets -----------------

#commented in the markdown as I've already done it
# st_write(pcdist2,"output_postcode_focus/postcode_focussed_data290621.gpkg")
# st_write(pcdist2,"output_postcode_focus/postcode_focussed_data290621.shp")
# 
# 
# pcdist2csv <- pcdist2
# st_geometry(pcdist2csv) <- NULL
# write_csv(pcdist2csv,"output_postcode_focus/postcode_focussed_data290621.csv")


```


## ANALYSIS OF DOGS PP AND CAR USE AT POSTCODE DISTRICT RESOLUTION (1% outiers removed)


The 99th percentile of dogs per person is as below.  22 rows were removed from the analysis in this appendix  

```{r}


######remove top 1% of dog ownership ###### 
###### this is in this appendix but not in the markdown describing the analysis in the paper#####
quantile(pcdist2$Dogs_pp,na.rm = TRUE,probs = c(0.99))
#0.9580356 

pcdist2 <- pcdist2 %>% filter(Dogs_pp <= 0.9580356 )
#removes 22 rows

```







I decided not to transform dogs_pp to log(dogs_pp) because keeping 
the original values makes interpretation easier. 
Using logs boosts the R squared slightly in the regression models,
 but the findings are the same: dog ownership remains a significant predictor even 
controlling for several different variables .

below is figure 1 from the paper and the spearman and pearson correlation overall i.e not by income quintile





```{r}


#####################################################################
##### ANALYSIS OF DOGS PP AND CAR USE AT POSTCODE DISTRICT RESOLUTION
#####################################################################

#---  dogs vs km pp --- by income quintile  

##I decided not to transform dogs_pp to log(dogs_pp) because keepint 
#the original values makes interpretation easier. 
#Using logs boosts the R squared slightly in the regression models,
# but the findings are the same: dog ownership remains a significant predictor even 
#controlling for several different variables 



require(hrbrthemes)
df <- pcdist2


#This plot is figure 1 in the paper 
df %>% filter(!is.na(meankm_pp)) %>%
  filter(!is.na(income_quintile)) %>%
  ggplot()+
  geom_point(aes(x = Dogs_pp, y = meankm_pp,colour=factor(income_quintile)))+
  geom_smooth(aes(x = Dogs_pp, y = meankm_pp),se = FALSE)+
  facet_wrap(~income_quintile)+
  #theme_ipsum() +
  scale_color_ipsum(name = "income quintile\n(1 = lowest)")+
  xlab("Dogs per person")+
  ylab("Car km travelled per person per year")

#correlations are high
cor(df$Dogs_pp,df$meankm_pp,use= "pairwise.complete.obs",method = "pearson")
#0.55
cor(df$Dogs_pp,df$meankm_pp,use= "pairwise.complete.obs",method = "spearman")
#0.667



```


The basic bivariate regression 
This is an interesting finding. Even without removing any outliers or transforming the predictor variable (dogs per person)  
An adjusted R squared of 0.3109  is achieved when dogs per person is the only 
predictor of car km travelled in England and Wales, 
 and it supports the hypothesis that dog ownership is a car dependent practice
 
*When outliers are removed the adjusted R squared is 0.3387  Slightly higher but similar * 
 

```{r}

r1 <- lm(data = df,meankm_pp~Dogs_pp)
summary(r1) #0.0.3109 

## This is an interesting finding. Even without removing any outliers or transforming the predictor variable (dogs per person)  
## An R squared of 00.3109  is achieved when dogs per person is the only 
## predictor of car km travelled in England and Wales, 
## and it supports the hypothesis that dog ownership is a car dependent practice.


library(broom)
reg_kmvsdog  <- tidy(r1)
#write_csv(reg_kmvsdog,"output_postcode_focus/reg_kmvsdog.csv")

```

Figure 2 Relationship between dog ownership and rural areas ]


when no outliers are removed (R = 0.63 Spearman)

*with outliers removed*  R =  `r  cor(df$Dogs_pp,df$meankm_pp,use= "pairwise.complete.obs",method = "spearman")` (spearman)


```{r}

#---------------------- plot Figure 2 Relationship between dog ownership and rural areas (R = 0.63 Spearman)---

require(hrbrthemes)
df <- pcdist2

#this plot makes figure 2 in the paper 
df %>% filter(!is.na(meankm_pp)) %>%
  filter(!is.na(ru5)) %>% #filter(!is.na(Dogs_pp)) %>%
  ggplot()+
  geom_point(aes(x = Dogs_pp, y = meankm_pp,colour=factor(ru5)))+
  geom_smooth(aes(x = Dogs_pp, y = meankm_pp),se = FALSE)+
  facet_wrap(~ru5)+
  #theme_ipsum() +
  scale_color_ipsum(name = "Urbanisation\n(1 = most urban)")+
  xlab("Dogs per person")+
  ylab("Car km travelled per person per year")




```


*correlations for figure 2 are very similar when outliers are removed.  The difference in R seems less than 0.01 for each subgroup*




```{r}

#correl;ations by RU5 
c <- df %>% filter(!is.na(meankm_pp)) %>%
  filter(ru5 ==1)
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "pearson") #0.46
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "spearman")#0.55

c <- df %>% filter(!is.na(meankm_pp)) %>%
  filter(ru5 ==2)
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "pearson") #0.22
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "spearman")#0.15

c <- df %>% filter(!is.na(meankm_pp)) %>%
  filter(ru5 ==3)
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "pearson") #0.41
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "spearman")#0.43

c <- df %>% filter(!is.na(meankm_pp)) %>%
  filter(ru5 ==4)
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "pearson") #0.10
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "spearman")#0.14

c <- df %>% filter(!is.na(meankm_pp)) %>%
  filter(ru5 ==5)
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "pearson") #0.11
cor(c$Dogs_pp,c$meankm_pp,use= "pairwise.complete.obs",method = "spearman")#-0.03

```


## Adding covariates e.g.  y = mean kmpp  vs dogs pp + other variables

the linear regression s are summarised below 

meankm_pp~ Dogs_pp + income_quintile + ru5

and then 

data = df,meankm_pp~ income_quintile + ru5

The model has a higher adjusted R squared when dogs are included.  

*This relationship is the same when outliers are removed* 

*All data*
with dogs adj R squared is 0.61 without dogs 0.58

*outliers removed* 
adj R squared is 0.61 without dogs 0.57


```{r}


############------------ Adding Covariates -----################################


#--------- adding covariates e.g.  y = mean kmpp  vs dogs pp + other variables

#This is presented in teh paper in table 2


#USE nolog in the paper because it's easier interpretation 

#km dogs income ru5 
r1 <- lm(data = df,meankm_pp~ Dogs_pp + income_quintile + ru5)
summary(r1) # # adj R qsuared is   0.6103
reg_kmvsDogIncomeRu5  <- tidy(r1)
#write_csv(reg_kmvsDogIncomeRu5 ,"output_postcode_focus/reg_kmvsDogIncomeRu5.csv")


#km income ru5 no dogs
r1 <- lm(data = df,meankm_pp~ income_quintile + ru5)
summary(r1) # # adj R qsuared is    0.5802 
reg_kmvsIncomeRu5  <- tidy(r1)
#write_csv(reg_kmvsIncomeRu5,"output_postcode_focus/reg_kmvsIncomeRu5.csv")


```



## The extended model 

#### Kent and Mulley's predictors 

First is a model just considering the factors considered in Kent and Mulley (2017), age, gender, dwelling type econ activity and car ownership.  this isn't the one in the paper.  the one in the paper has these Kent and Mulley variables in addition to the urbanisation and income variables which we've already considered.  



```{r}


#--------- model the types of variables considered by Kent and Mulley ------

# there wasn't space to include this step in the paper

#we are going to consider a number of variables with which there is correlation 

#let's make a correlation plot of the variables we will consider in the model

df <- pcdist2 %>% select(PostDist,meankm_pp,Dogs_pp,age_median,dwel_detached_percent,
                         dwel_semi_detached_percent,dwel_terraced_percent,flat_percent,
                         econ_all_econ_active_percent,
                         nocar_percent,pc_female)


kent_mulley_data <- df 
st_geometry(kent_mulley_data) <- NULL
#write_csv(kent_mulley_data ,"output_postcode_focus/kent_mulley_data.csv")

#library(corrplot)
#library(Hmisc)
#corrplot(df, method = "shade",type = "upper", order = "hclust", 
#         tl.col = "black", tl.srt = 45)


#install.packages("ggcorrplot")
require(ggcorrplot)

corr <- df 
st_geometry(corr)<- NULL
corr <- corr %>% drop_na() %>% select(-PostDist)


corr <- round(cor(corr), 1)
ggcorrplot(corr, hc.order = TRUE, type = "lower",
           outline.col = "white",
           ggtheme = ggplot2::theme_gray,
           colors = c("#6D9EC1", "white", "#E46726"),lab = TRUE)



#names(pcdist2)




```

Predicting km per person without dogs

```{r}

#######---- without removing any outliers #######------------- 

#no logs for ease of interpretation 
## without dogs
r1 <- lm(data = df,meankm_pp~ age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent+
           nocar_percent
)

summary(r1) #Adjusted R-squared:  0.8324  
kent_mulley_nodog  <- tidy(r1)
#write_csv(kent_mulley_nodog ,"output_postcode_focus/kent_mulley_nodog.csv")

```

Predicting km per person with dogs

```{r}
#this is the model to try to predict car km travelled if we include dog ownership 
r1 <- lm(data = df,meankm_pp~ Dogs_pp + age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent+
           nocar_percent
           )

summary(r1) #Adjusted R-squared:  0.8403  
kent_mulley_withdog  <- tidy(r1)
#write_csv(kent_mulley_withdog ,"output_postcode_focus/kent_mulley_withdog.csv")

```




####  The extended model (D and E ) used in the paper with outliers removed


*This relationship still holds when outliers are removed: including dogs pp in the model improves the Adj R squared suggesting dog ownership is associated with car use after controlling for other factors* 

*All data*
with dogs adj R squared is 0.8752  with dogs and  0.8718 without dogs

*outliers removed* 
with dogs adj R squared is 0.8717  with dogs and  0.8745 without dogs



```{r}

#--------------------- kent and mulley + income and urbanisation  -----

##this overall combined model is included in the paper (models D & E)


#we are going to consider a number of variables with which there is correlation 

#let's make a correlation plot of the variables we will consider in the model

df <- pcdist2 %>% select(PostDist,meankm_pp,Dogs_pp,income_quintile,ru5,age_median,dwel_detached_percent,
                         dwel_semi_detached_percent,dwel_terraced_percent,flat_percent,
                         econ_all_econ_active_percent,
                         nocar_percent,pc_female)


kent_mulley_data <- df 
st_geometry(kent_mulley_data) <- NULL
#write_csv(kent_mulley_data ,"output_postcode_focus/kent_mulley_income_ru5_data.csv")



```


Without dogs per person as a predictor


```{r}

#######---- without removing any outliers #######------------- 

#no logs for ease of interpretation 
## without dogs
r1 <- lm(data = df,meankm_pp~ income_quintile +ru5 +  age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent+
           nocar_percent
)

summary(r1) #Adjusted R-squared:  0.8718  
kent_mulley_nodog  <- tidy(r1)
#write_csv(kent_mulley_nodog ,"output_postcode_focus/kent_mulley_income_ru5_nodog.csv")



```


With dogs per person as a predictor The Adjusted R squared is still slightly higher than without it even when a large number of other predictors are included in the model.  



```{r}

#this is the model to try to predict car km travelled if we include dog ownership 
r1 <- lm(data = df,meankm_pp~ Dogs_pp +income_quintile +ru5 + age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent+
           nocar_percent
)

summary(r1) #Adjusted R-squared:  0.8752  
kent_mulley_withdog  <- tidy(r1)
#write_csv(kent_mulley_withdog ,"output_postcode_focus/kent_mulley_income_ru5_withdog.csv")
```



This is the end of the analysis where we remove some outliers to check if the relationship still holds.




## Examining the effect of taking logs 


We could transform the predictor dogs per person. Avoiding transformation allows easier interpretation, but we should check that the findings hold if we take logs 




#### Plotting the graph used in figure 1 with log(dogs per person) and outliers removed 



The plot and correlations with 1% outliers removed AND log(dogs per person).  In this the difference between pearson ans spearman correlation coefficients is smaller.  




```{r}

require(hrbrthemes)
df <- pcdist2


#This plot is figure 1 in the paper 
df %>% filter(!is.na(meankm_pp)) %>%
  filter(!is.na(income_quintile)) %>%
  ggplot()+
  geom_point(aes(x = log(Dogs_pp), y = meankm_pp,colour=factor(income_quintile)))+
  geom_smooth(aes(x = log(Dogs_pp), y = meankm_pp),se = FALSE)+
  facet_wrap(~income_quintile)+
  #theme_ipsum() +
  scale_color_ipsum(name = "income quintile\n(1 = lowest)")+
  xlab("log(Dogs per person)")+
  ylab("Car km travelled per person per year")

#correlations are high
cor(log(df$Dogs_pp),df$meankm_pp,use= "pairwise.complete.obs",method = "pearson")
#0.55
cor(log(df$Dogs_pp),df$meankm_pp,use= "pairwise.complete.obs",method = "spearman")
#0.667


```


####  The extended model (D and E ) used in the paper with outliers removed AND log(dogs per person)


*This relationship still holds when outliers are removed: including dogs pp in the model improves the Adj R squared suggesting dog ownership is associated with car use after controlling for other factors* 

*All data*
with dogs adj R squared is 0.8752  with dogs and  0.8718 without dogs

*outliers removed* 
with dogs adj R squared is 0.8717  with dogs and  0.8745 without dogs


The relationship is still stronger with log(dogs per person) as a covariate.  

*outliers removed and log(dogs per person)*

with dogs adj R squared is 0.8777  with dogs and  0.8745 without dogs

```{r}

#--------------------- kent and mulley + income and urbanisation  -----

##this overall combined model is included in the paper (models D & E)


#we are going to consider a number of variables with which there is correlation 

#let's make a correlation plot of the variables we will consider in the model

df <- pcdist2 %>% select(PostDist,meankm_pp,Dogs_pp,income_quintile,ru5,age_median,dwel_detached_percent,
                         dwel_semi_detached_percent,dwel_terraced_percent,flat_percent,
                         econ_all_econ_active_percent,
                         nocar_percent,pc_female)


kent_mulley_data <- df 
st_geometry(kent_mulley_data) <- NULL
#write_csv(kent_mulley_data ,"output_postcode_focus/kent_mulley_income_ru5_data.csv")



```


Without dogs per person as a predictor


```{r}

#######---- without removing any outliers #######------------- 

#no logs for ease of interpretation 
## without dogs
r1 <- lm(data = df,meankm_pp~ income_quintile +ru5 +  age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent+
           nocar_percent
)

summary(r1) #Adjusted R-squared:  0.8718  
kent_mulley_nodog  <- tidy(r1)
#write_csv(kent_mulley_nodog ,"output_postcode_focus/kent_mulley_income_ru5_nodog.csv")



```


With dogs per person as a predictor The Adjusted R squared is still slightly higher than without it even when a large number of other predictors are included in the model.  




```{r}

#this is the model to try to predict car km travelled if we include dog ownership 
r1 <- lm(data = df,meankm_pp~ log(Dogs_pp) +income_quintile +ru5 + age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent+
           nocar_percent
)

summary(r1) #Adjusted R-squared:  0.8752 (original model in paper) 
kent_mulley_withdog  <- tidy(r1)
#write_csv(kent_mulley_withdog ,"output_postcode_focus/kent_mulley_income_ru5_withdog.csv")
```





##  If we do not include car ownership in models D and E 

If we do not include car ownership in models D and E, the adjusted R squared is still slightly higher with dogs than without.  NB This check on the analysis does not use logs of dogs per person.  

no dogs 
Adjusted R-squared:  0.8718 with cars   0.8639 without


with dogs: 
Adjusted R-squared:  0.8752 with car ownership Adjusted R-squared:  0.8678 with car ownership removed



```{r}



### --------- what if we don't include car ownership ?? 


df <- pcdist2 %>% select(PostDist,meankm_pp,Dogs_pp,income_quintile,ru5,age_median,dwel_detached_percent,
                         dwel_semi_detached_percent,dwel_terraced_percent,flat_percent,
                         econ_all_econ_active_percent,
                         pc_female)


kent_mulley_data <- df 
st_geometry(kent_mulley_data) <- NULL
#write_csv(kent_mulley_data ,"output_postcode_focus/kent_mulley_income_ru5_data.csv")





#no logs for ease of interpretation 
## without dogs
r1 <- lm(data = df,meankm_pp~ income_quintile +ru5 +  age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent
           
)

summary(r1) #Adjusted R-squared:  0.8718 with cars   0.8639 without
kent_mulley_nodog  <- tidy(r1)
#write_csv(kent_mulley_nodog ,"output_postcode_focus/kent_mulley_income_ru5_nodog.csv")


#with dogs but without car ownership 
r1 <- lm(data = df,meankm_pp~ Dogs_pp +income_quintile +ru5 + age_median + pc_female + dwel_detached_percent +
           dwel_semi_detached_percent + dwel_terraced_percent + flat_percent+
           econ_all_econ_active_percent
           
)

summary(r1) #Adjusted R-squared:  0.8752 with car ownership Adjusted R-squared:  0.8678 with car ownership removed 
kent_mulley_withdog  <- tidy(r1)





```






